---
title: "STAT_626 Assignment 04"
author: "Philip Anderson; panders2@tamu.edu"
date: "6/14/2018"
output: pdf_document
---

```{r setup, message=FALSE}
library("astsa")
library("tidyverse")
```

```{r, echo=F}
#ARMAtoMA(ar=0.5, ma=0, 10)
```

```{r, echo=F}
#ARMAtoMA(ar=c(1, -0.5), ma=-1, 10)
```

# Question 3.4
x_t = -0.9x_t-2 + w_t

```{r}
# give the roots of the characteristic equation
z <- c(1, 0, .9)
# solve for the roots of autoregressive polynomial
paste0("Roots of Autoregressive Polynomial given by: " ,(a <- polyroot(z)[1]))
arg <- Arg(a)/(2*pi)
paste0("Pseudo-period given by: ", 1 / arg) # pseudo-period

```


```{r}
set.seed(1914)
ar2 = arima.sim(list(order=c(2,0,0), ar=c(0,-.9)), n = 48)
plot(ar2, axes=FALSE, xlab="Time",main="Simulated Values from Specified AR(2) model")
axis(2)  
axis(1, at=seq(0,48,by=4))  
box()
abline(v=seq(0,48,by=4), lty=2)


```


```{r}
# plot the ACF
ACF = ARMAacf(ar=c(0,-.9), ma=0, 50)
plot(ACF, type="h", xlab="lag", main="Theoretical ACF for AR(2) model")
abline(h=0)
```

Because the coefficient on the AR model is negative, the lags will flip between being positive and negative at every lag iteration.

# Question 3.17
## Question 3.17.A

```{r}
astsa::tsplot(cpg ,main="Plot of c_t")
```

The plot seems to be showing an exponential decrease in the price per gigabyte of storage.

## Question 3.17.B

```{r}
# regress log(c_t) on a time index
lin_mod <- lm(log(cpg) ~ time(cpg))
```


```{r}
# plot the fitted line to compare it to the logged data
ts.plot(log(cpg))
abline(lin_mod)
```

The fitted regression line is very close to the log transformation of the data.  $$c_t \approx \alpha exp(\beta t) $$ 

## Question 3.17.C
Plot the residuals of the model and comment.

```{r}
plot(density(resid(lin_mod)), main="Density Plot of Residuals")
```

The density plot looks good, but keep in mind that we may have serial correlation that this isn't capturing.


```{r}
ts.plot(resid(lin_mod), main="Residual Time Series Plot")
acf2(resid(lin_mod))
```

It appears from the time series plot and the ACF of the residuals that the errors display some level of serial or autocorrelation.  A regression that does not assume the errors to be i.i.d. may be necessary.

## Question 3.17.D

```{r}
# refit the model, but using the fact that the errors are autocorrelated
serial_mod <- sarima(
                    xdata=log(cpg)
                    , p=c(1)
                    , d=c(0)
                    , q=c(0)
                    , xreg=c(time(cpg))
                    )

```

```{r}
astsa::tsplot(serial_mod$fit$residuals, main="Time series Plot of AutoCorrelated Regression")
acf2(serial_mod$fit$residuals)
```

This appears to be a better fit to the data - we have eliminated the evidence of serial correlation from the model errors.

```{r}
paste0("Coefficients for new Model: ")
print(serial_mod$fit$coef)

cat("\nCoefficients for Linear Model: ")
print(lin_mod$coefficients)

```