---
title: "Assignment 05; STAT 626"
author: "Philip Anderson; panders2@tamu.edu"
date: "7/03/2018"
output: pdf_document
---

```{r setup, include=FALSE}
library("astsa")
```

# Question 3.19
Plot theoretical ACF of SARIMA(0, 0, 1) x (1,0,0)12.  
PHI = 0.8, theta=0.5. lags=50

```{r}
(phi <- c(rep(0, 11), 0.8))
paste0("phi length: ", length(phi))

ACF <- ARMAacf(ar=phi, ma=0.5, lag.max=50)
plot(ACF, type="h"
     , xlab="lags" 
     , main=expression(paste("Theoretical ACF: ", Phi[12], "=0.8 ", theta, "=0.5")))
```


# Question 3.20

```{r}
# first plot the 
par(mfrow=c(4,1))
astsa::tsplot(chicken, main="Time-Index Plot of Chicken")
astsa::tsplot(diff(chicken), main="First Difference Taken")
astsa::tsplot(diff(diff(chicken), 4), main="Quarterly Difference of First Difference")
astsa::tsplot(diff(diff(chicken), 12), main="Annual Difference of First Difference")
```

```{r, echo=F,  results='hide'}
acf2(diff(chicken), max.lag=100)
acf2(diff(diff(chicken), 4), max.lag=100)
acf2(diff(diff(chicken), 12), max.lag=100)
```

From the above ACFs+PACFs, it appears that annual differencing is a good option.  The final ACF+PACF combination shows a tailing off in the ACF, but a cut-off in the PACF after seasonal lag 3.  This suggests that a suitable model may be SAR(3), P=3, Q=0 in S=12, or SARIMA(3,1,0)x(3,1,0)_12.

```{r, echo=F, results='hide'}
smod <- sarima(xdata=chicken, p=3, d=1, q=0, P=3, D=1, Q=0, S=12)
print(smod)
```

The above plots look ok, but the Ljung-Box p-values are highly significant, and the model fit procedure failed, which is not a good sign.  Let's loop through some potential variations of the model and see if we can improve on the BIC.

I am going to do a nested for-loop.  AR differences of [1,2,3], MA differences of [1,2,3], at both the immediate and seasonal levels.  Also going to try Seasonal = 4, 12, in case the chicken prices fluctuate quarterly.


```{r, fig.keep='none'}
puzzle <- cbind(0, 0, 0, 0, 0)

for (k in c(12, 4)) {
  for (i in 0:3) {
    for (j in 0:3) {
      smod <- try(sarima(xdata=chicken, p=i, d=1, q=j, P=i, D=1, Q=j, S=k), silent=T)
      
        if(class(smod)=="list") {
            piece <- cbind(i, j, k, smod$AIC, smod$BIC)
                                }
        else {
            piece <- cbind(i, j, k, 0, 0)
             }

      puzzle <- rbind(puzzle, piece)
    }
  }
}
```

```{r}
puzzle
```

Of the models evaluated, it looks like SARIMA(3,1,1)x(3,1,1)12 is the strongest combination of what I had originally hypothesized and what is borne out by fit statistics.

```{r}
sarima(xdata=chicken, p=3, d=1, q=1, P=3, D=1, Q=1, S=12)

```

```{r}
sarima.for(xdata=chicken, p=3, d=1, q=1, P=3, D=1, Q=1, S=12, n.ahead=12)
```


# Question 3.21

```{r}
par(mfrow=c(4,1))
astsa::tsplot(UnempRate)
astsa::tsplot(log(UnempRate))
astsa::tsplot(diff(log(UnempRate)))
astsa::tsplot(diff(diff(log(UnempRate)), 12))

```

```{r}
acf2(diff(diff(log(UnempRate)), 12), 75)
```

The ACF cuts off after the first season, but the PACF tails off.  This could suggest a SAM model of (0,1,1)x(0,1,1)12, but we will work through several iterations to make sure.


```{r, fig.keep='none'}
puzzle <- cbind(0, 0, 0, 0, 0)

for (k in c(12)) {
  for (i in 0:3) {
    for (j in 0:3) {
      smod <- try(sarima(xdata=UnempRate, p=i, d=1, q=j, P=i, D=1, Q=j, S=k), silent=T)
      
        if(class(smod)=="list") {
            piece <- cbind(i, j, k, smod$AIC, smod$BIC)
                                }
        else {
            piece <- cbind(i, j, k, 0, 0)
             }

      puzzle <- rbind(puzzle, piece)
    }
  }
}
```


```{r}
puzzle
```

The model hypothesized does not have dramatically different results from the others.  

```{r}

smod <- sarima(xdata=UnempRate, p=0, d=1, q=1, P=0, D=1, Q=1, S=12)
smod
```

```{r}
sarima.for(xdata=UnempRate, p=0, d=1, q=1, P=0, D=1, Q=1, S=12, n.ahead=12)
```


# Question 3.24

Rerun the model in that example but fit an ARIMA(2, 0, 0) × (0, 1, 1)12 to the residuals. Does this improve the fit?

```{r}
# replicate the original example
dummy <- ifelse(soi<0, 0, 1)
fish  = ts.intersect(rec, soiL6=lag(soi,-6), dL6=lag(dummy,-6), dframe=TRUE)
summary(fit <- lm(rec ~soiL6*dL6, data=fish, na.action=NULL)) 
tsplot(resid(fit))
acf2(resid(fit))
intract <- fish$soiL6*fish$dL6
sarima(xdata=rec, p=2, d=0, q=0, xreg=cbind(fish$soiL6, fish$dL6, intract))

```

```{r}
# now, rerun the SARIMA model with seasonal component
sarima(xdata=rec, p=2, d=0, q=0, P=0, D=1, Q=1, S=12, xreg=cbind(fish$soiL6, fish$dL6, intract))
```

The specified SARIMA model is a marginally better fit over the ARIMA model.


