---
title: "panders2_stat626_hw03"
author: "Philip Anderson; panders2@tamu.edu"
date: "6/6/2018"
output: pdf_document
---

```{r setup, message=FALSE}
library("tidyverse")
library("gmodels")
```


# Question I
## Question I) a)

```{r}
# simulate two random walks with initial values = 0

set.seed(7)
y <- cumsum(c(0, rnorm(99, 0, 1)))
set.seed(8)
x <- cumsum(c(0, rnorm(99, 0, 1)))

```

## Question I) a) i)

```{r}

plot(x, y, pch=16, cex=1.25
     , main="x realization vs. y realization ScatterPlot"
     )

```

From the above, it appears that there is a non-linear negative relationship between y and x.  

```{r}
# nonparametric correlation coefficient
cor(x, y, method="spearman")
```

## Question I) a) ii)

I expect the hypothesis to be rejected.  There is a clear (to me) negative relationship between the variables, and we have a non-trivial amount of data.

## Question I) a) iii)

```{r}
lin_mod <- lm(y ~ x)
summary(lin_mod)
```

From the summary printout, it is clear the beta_1 coefficient is negative, and we are soundly rejecting the null hypothesis.  Because of the random nature of the random walk, I would not necessarily expect this relationship to hold under repeated trials.

## Question I) b)

```{r}

beta_val_vec <- numeric(1000)
p_val_vec <- numeric(1000)

for (i in 1:1000) {

      set.seed(i)
      y <- cumsum(c(0, rnorm(99, 0, 1)))
      set.seed(5000 + i)
      x <- cumsum(c(0, rnorm(99, 0, 1)))
      
      lin <- lm(y ~ x)
      lin_summ <- summary(lin)
      beta_val <- lin_summ$coefficients[2,1]
      p_val <- lin_summ$coefficients[2,4]
      beta_val_vec[i] <- beta_val
      p_val_vec[i] <- p_val
      
}

outcome <- data.frame(cbind(beta_val_vec, p_val_vec))
```

```{r}


outcome$beta_pos <- ifelse(outcome$beta_val_vec >= 0, "pos", "neg")
outcome$signif <- ifelse(outcome$p_val_vec < 0.05, "sig", "not_sig")

table(outcome$beta_pos, outcome$signif)
```

```{r}
gmodels::CrossTable(outcome$beta_pos, outcome$signif)
```

```{r}

length(p_val_vec[p_val_vec > 0.05]) / length(p_val_vec)

length(beta_val_vec[beta_val_vec > 0])
length(beta_val_vec[beta_val_vec < 0])

```





